\section*{Acknowledgements}

I approached Prof. Schmid with the idea of writing this thesis building upon his previous research \citetitle{zabka_empirical_2022} ~\cite{zabka_empirical_2022} and \citetitle{zabka_short_2022} ~\cite{zabka_short_2022}.
Since the very start he was supporting me to realize this project: \tech{ln-history}.
With his help, I got in touch with Dr. Christian Decker who generously sent to me his collected gossip data and 
also assisted me tremendously with his MIT licensed code found in the \textit{lnresearch} GitHub repository~\cite{decker_lnresearch_2020}.
After exploring Lightning gossip messages and my growing interest in graph theory,
I delved into the fundamentals of the Lightning Network and started building tools around the gossip data.

Thanks to the Socratic Seminar run by the BitDevs Berlin, I got in touch with the renowned Bitcoin Core Developer Fabian Jahr
who proactively arranged a meeting with the highly respected Lightning researcher René Pickhardt.

Around that time Chaincode Labs hosted its summer school \term{Summer of Bitcoin}\footnote{\url{https://www.summerofbitcoin.org/}} where I applied
with the idea of implementing \tech{ln-history} and to which I got accepted.
René and I worked together in a span of three months which turned out to be an incredibly meaningful experience for me.
With a background in full-stack software engineering I was quickly humbled by the never ending challenges and problems.
René as my mentor did not only help me to stay on track from a technical and engineering point of view 
but also from an organizational which in retro perspective enabled me to proudly release \tech{ln-history} \code{v1.0}\footnote{\url{https://ln-history.info}}.

In October 2025 the lightning++ conference happened in Berlin where I got in touch with various Lightning developers.
Some of them gave me their collected gossip messages, which made the \tech{ln-history} gossip data set grow to \num{91.3} million messages.
Additionally, I gathered feedback, new ideas and the opportunity to present the first findings at the Bitcoin Forum Bayern 2025.

I also want to thank my friends, family and girlfriend who have proofread this thesis.
Lastly, I want to thank every open-source contributor who altruistically works on software that shapes our world.